sensiAnPaper1All

### 2015-07-30 sensiAnPaper1All.Rmd
improve tree pruning by allowing for error

@@

  #hardcode which variables to include in the tree analysis here
  #to keep it simple and transparent
  treePredictors <- c('P_1','P_2','startResistanceRatio','exposure','phi.SS1_A0','phi.SS2_0B','h.RS1_A0','h.RS2_0B','s.RR1_A0','s.RR2_0B')
  treeInput <- treeInput[ treePredictors, ]

@@

### do tree analysis
```{r, eval=FALSE}
  require(rpart.plot)

  #transpose
  treeInput <- t(treeInput)
  #convert to a dataframe
  treeInput <- data.frame(treeInput)
  #create string with predictor names in
  treePredictorString <- paste(treePredictors, collapse="+")

@@

    tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class') 
    treePruned <- prune(tree, cp=tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"])
    #adds in no. scenarios correctly and incorrectly classified
    #prp(tree, extra=1, varlen = 10, main=treeResponse)
    prp(treePruned, extra=1, varlen = 10, main=treeResponse)

@@

    #trying to limit num levels a different way using maxdepth
    tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class', maxdepth=3) 
    #adds in no. scenarios correctly and incorrectly classified
    prp(tree, extra=1, varlen = 10, main=treeResponse)
    savePlot(paste0(outFolder,"tree",treeResponse,"depth3.jpg"), type='jpg')

  }

#   treeResponse <- 'betterMix1Seq_cP0.5'
#   tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class') 
#   
#   #if get error number of rows of matrices must match (see arg 2)
#   #is because all vals are 0
# 
#   plot(tree) #plots tree
#   text(tree) #labels tree
#   #branch predictors 2 phis & P_2  
#     
#   #for mix3 (resistance to both)
#   x11()
#   treeResponse <- 'betterMix3Seq_cP0.5'
#   tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class')  
#   plot(tree) #plots tree
#   text(tree) #labels tree  
#   
#   #branch predictors 2 phis & P_1
#   
#   #or from rpart.plot
#   #prp(tree)
#   
#   #adds in no. scenarios correctly and incorrectly classified
#   prp(tree, extra=1, varlen = 10, main=treeResponse)
#   savePlot(paste0(outFolder,"tree",treeResponse,".png"), type='png')

@@

    #susanas code
    #pruning tree
    #   treePruned <- prune(tree, cp=tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"])
    #   plot(treePruned) #plots tree
    #   text(treePruned) #labels tree 
    # prp(prunned_tree, extra=1,box.col=c("pink", "palegreen3")[prunned_tree$frame$yval],cex=0.5,yesno.yshift=-2)

>>>>

  #renaming other rownames to make nicer plots
  rownames(treeInput)[rownames(treeInput)=="phi.SS1_A0"] <- "effectiveness_insecticide1"
  rownames(treeInput)[rownames(treeInput)=="phi.SS2_0B"] <- "effectiveness_insecticide2"

@@

  #hardcode which variables to include in the tree analysis here
  #to keep it simple and transparent
  #treePredictors <- c('P_1','P_2','startResistanceRatio','exposure','phi.SS1_A0','phi.SS2_0B','h.RS1_A0','h.RS2_0B','s.RR1_A0','s.RR2_0B')
  #prettifying names
  treePredictors <- c('P_1','P_2','startResistanceRatio','exposure','effectiveness_insecticide1','effectiveness_insecticide2','h.RS1_A0','h.RS2_0B','s.RR1_A0','s.RR2_0B')
  treeInput <- treeInput[ treePredictors, ]

  #replace 0s & 1s with mixture/sequence to make plot clearer
  repMix <- function(x){
    x[x==0] <- 'sequence'
    x[x==1] <- 'mixture'
    x
  }

  resistBetterMix1Seq <- repMix(resistBetterMix1Seq)
  resistBetterMix2Seq <- repMix(resistBetterMix2Seq)
  resistBetterMix3Seq <- repMix(resistBetterMix3Seq)

  resistBetterMix1Seq20 <- repMix(resistBetterMix1Seq20)
  resistBetterMix2Seq20 <- repMix(resistBetterMix2Seq20)
  resistBetterMix3Seq20 <- repMix(resistBetterMix3Seq20)

@@

### do tree analysis
```{r, eval=TRUE}
  require(rpart.plot)

  #transpose
  treeInput <- t(treeInput)

  #convert to a dataframe
  treeInput <- data.frame(treeInput, stringsAsFactors = FALSE)

  #convert predictor columns back to numeric
  #following this above : replace 0s & 1s with mixture/sequence to make plot clearer
  filter <- !substr(names(treeInput),1,6) %in% 'better'
  treeInput[,filter] <- lapply(treeInput[,filter],as.numeric) 
  filter <- !filter
  treeInput[,filter] <- lapply(treeInput[,filter],as.factor)

  #create string with predictor names in
  treePredictorString <- paste(treePredictors, collapse="+")

@@

    tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class') 
    #treePruned <- prune(tree, cp=tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"])

    #http://stackoverflow.com/questions/29197213/what-is-the-difference-between-rel-error-and-x-error-in-a-rpart-decision-tree
    #A rule of thumb is to choose the lowest level where the rel_error + xstd < xerror.

    cpOptimal <- tree$cptable[ tree$cptable[,"rel error"] + tree$cptable[,"xstd"]  < tree$cptable[,"xerror"],"CP"][1]

    treePruned <- prune(tree, cp=cpOptimal)

    #to set box colours
    cols <- ifelse(treePruned$frame$yval == 1, "red", "green3")

    #adds in no. scenarios correctly and incorrectly classified
    #prp(tree, extra=1, varlen = 10, main=treeResponse)
    #prp(treePruned, extra=1, varlen = 10, main=treeResponse, col=cols, border.col=cols)
    #prp(treePruned, extra=1, varlen = 0, main=treeResponse, col=cols, border.col=cols)
    prp(treePruned, extra=1, varlen = 0, main=treeResponse, box.col=cols, under=TRUE)

@@

    #trying to limit num levels a different way using maxdepth
    #tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class', maxdepth=3) 
    #adds in no. scenarios correctly and incorrectly classified
    #prp(tree, extra=1, varlen = 10, main=treeResponse)
    #savePlot(paste0(outFolder,"tree",treeResponse,"depth3.jpg"), type='jpg')

  }
```

### testing tree analysis for paper
```{r, eval=TRUE}

    x11()

    treeResponse <- "betterMix2Seq_cP0.5"

    tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class') 
    treePruned <- prune(tree, cp=tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"])

    #to set box colours
    cols <- ifelse(treePruned$frame$yval == 1, "red", "green3")

    #adds in no. scenarios correctly and incorrectly classified
    #prp(treePruned, extra=1, varlen = 0, main=treeResponse, box.col=cols, under=TRUE)

    cpOptimal <- tree$cptable[ tree$cptable[,"rel error"] + tree$cptable[,"xstd"]  < tree$cptable[,"xerror"],"CP"][1]

    treePruned <- prune(tree, cp=cpOptimal)

    prp(treePruned, extra=1, varlen = 0, main=treeResponse, box.col=cols, under=TRUE)

    #comparing it with 0/1 gives the same big tree
    treeInput$betterMix2Seq_cP0.5numeric <- as.numeric(treeInput$betterMix2Seq_cP0.5)

    treeResponse <- "betterMix2Seq_cP0.5numeric"

    tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class') 
    treePruned <- prune(tree, cp=tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"])

    #to set box colours
    cols <- ifelse(treePruned$frame$yval == 1, "red", "green3")

    prp(treePruned, extra=1, varlen = 0, main=treeResponse, box.col=cols, under=TRUE)

    #testing alternative pruning
    #A rule of thumb is to choose the lowest level where the rel_error + xstd < xerror.
    treePruned <- prune(tree, cp=tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"])

```    
#   treeResponse <- 'betterMix1Seq_cP0.5'
#   tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class') 
#   
#   #if get error number of rows of matrices must match (see arg 2)
#   #is because all vals are 0
# 
#   plot(tree) #plots tree
#   text(tree) #labels tree
#   #branch predictors 2 phis & P_2  
#     
#   #for mix3 (resistance to both)
#   x11()
#   treeResponse <- 'betterMix3Seq_cP0.5'
#   tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class')  
#   plot(tree) #plots tree
#   text(tree) #labels tree  
#   
#   #branch predictors 2 phis & P_1
#   
#   #or from rpart.plot
#   #prp(tree)
#   
#   #adds in no. scenarios correctly and incorrectly classified
#   prp(tree, extra=1, varlen = 10, main=treeResponse)
#   savePlot(paste0(outFolder,"tree",treeResponse,".png"), type='png')

@@

    #susanas code
    #pruning tree
    #   treePruned <- prune(tree, cp=tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"])
    #   plot(treePruned) #plots tree
    #   text(treePruned) #labels tree 
    # prp(prunned_tree, extra=1,box.col=c("pink", "palegreen3")[prunned_tree$frame$yval],cex=0.5,yesno.yshift=-2)
```
